---
title: "Assignment B-1"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## 1. Introduction

This mini-project is part of the assignments in the STAT545B course. It involves the following exercises:

1. Making a function and fortifying it 
2. Documenting the function 
3. Demonstrating the usefulness of the function with examples
4. Testing the function
5. Making a README file for the project repository

Ultimately, a well written, reproducible Rmd file containing all the code used in the exercises and relevant descriptions will be knitted to a markdown file and pushed to the new Github repository created as part of the course.


### 1.1. Load required packages

The packages below will be used to complete the exercises in this project. Also, the
`palmerpenguins::penguins`, `gapminder::gapminder`, `datateachr::apt_buildings` datasets will be used. Therefore, I will load them before I continue with the exercises. 

However, if any of the packages is not installed, the function `install.packages()` should be used for that purpose before loading.

```{r}
suppressMessages(library(tidyverse))
suppressMessages(library(palmerpenguins))
suppressMessages(library(gapminder))
suppressMessages(library(testthat))
suppressMessages(library(datateachr))
suppressMessages(library(roxygen2))
```


## 2. Exercises 1 and 2: Make a function and document it

Here, I will make, fortify and document a function that takes a tibble or data frame, computes summary statistics on a numeric variable grouped by a categorical variable, and outputs a tibble or data frame and a boxplot.

```{r}
# Document function using the roxygen2 package

#' Compute summary statistics, produce a tibble or data frame and a boxplot
#' 
#' This function computes summary statistics on a numeric variable 
#' grouped by a categorical variable. The summary statistics include 
#' mean, median, minimum, maximum, and count. It also 
#' produces a boxplot.
#' 
#' @param df Tibble or data frame containing variables for computing summary statistics. 
#' Name is easily recognizable.
#' @param x A categorical variable by which rows are grouped. 
#' It is named x because it is more generally used and easy to identify.
#' @param y A numeric variable for computing summary statistics. 
#' It is named y because it is more generally used and easy to identify.
#' @return A tibble or data frame containing summary statistics and a boxplot.


summarize_data <- function(df, x, y, na.rm = TRUE) {
    val1 <- eval(substitute(x), df)
    val2 <- eval(substitute(y), df)

    if (!(is.character(val1) || is.factor(val1))) {
      stop('This function only works with a categorical input as x.\n',
           'You have provided an object of class: ', class(val1))
    }

    if (!is.numeric(val2)) {
      stop("This function only works with a numeric input as y.\n",
           "You have provided a variable of class: ", class(val2))
    }
  
    summary_df <- df %>%
    group_by({{ x }}) %>%
    summarise(mean = mean({{ y }}, na.rm = na.rm),
              median = median({{ y }}, na.rm = na.rm),
              min = min({{ y }}, na.rm = na.rm),
              max = max({{ y }}, na.rm = na.rm),
              n = n())
  
    summary_plot <- df %>% ggplot(aes({{ x }}, {{ y }})) +
    geom_boxplot()
  
    return(list(summary_df, summary_plot))
}

```

The new function takes a tibble or data frame, removes all missing values by default (user can specify otherwise), groups the data by a categorical variable, and calculates mean, median, minimum, maximum, and count of a numeric variable. It returns a tibble or data frame and a boxplot.

To fortify the function, I generalized it, included the option to intentionally allow missing values by changing the default from `na.rm = TRUE` to `na.rm = FALSE`. I also documented the function.

With the code established above, I can easily pass the new function `summarize_data` to other datasets that meet the criteria set in the code.


## 3. Exercise 3: Include examples

I will use the following examples to demonstrate the usage of the `summarize_data` function.

```{r}
# Example 1: Compute summary statistics using the palmerpenguins::penguins data

summarize_data(penguins, species, bill_length_mm)

```

The function call above groups the penguins data by `species`, computes summary statistics on `bill_length_mm`, produces a tibble and a boxplot.



```{r}
# Example 2: Compute summary statistics using the gapminder::gapminder data

summarize_data(gapminder, continent, lifeExp)
```

The function call above groups the `gapminder` data by `continent`, computes summary statistics on `lifeExp`, produces a tibble and a boxplot.



**Now, let me use a numeric variable in place of a categorical variable for x and see what the outcome would be.**

```{r error=TRUE}

summarize_data(gapminder, gdpPercap, lifeExp)
```

The function throws an error, since a numeric variable was used in place of a categorical variable.


**I will also use a categorical variable where a numeric variable is expected and see if the function would throw an error**

```{r error = TRUE}
summarize_data(gapminder, continent, country)
```


## 4. Exercise 4: Test the function

I will use the `expect_()` function in the `test_that` package to test the newly created `summarize_data()` function

```{r}
# First, compute summary statistics manually
summary1 <- penguins %>% 
  group_by(species) %>%
  summarize(mean = mean(body_mass_g, na.rm = TRUE))
# Extract the mean column
  summary1$mean

  
# Compute summary statistics using the summarize_data function
summary2 <- summarize_data(penguins, species, body_mass_g)
# Extract the tibble, then the mean column
  summary2[[1]]$mean

test_that("the two means are identical", {
  expect_identical(summary1$mean, summary2[[1]]$mean)})

test_that("the function returns a list", {
  expect_type(summarize_data(apt_buildings, window_type, no_of_storeys), "list")})
  
test_that("the use of a numeric variable for x returns an error", {
  expect_error(summarize_data(apt_buildings, year_built, no_of_storeys))})

test_that("the two mean vectors have NAs", {
  na_present_1 <- penguins %>% 
                  group_by(species) %>% 
                  summarise(mean = mean(body_mass_g))
  na_present_2 <- summarize_data(penguins, species, body_mass_g, na.rm = FALSE)
  expect_equal(na_present_1$mean, na_present_2[[1]]$mean)})
  
test_that("silent when function successfully creates a boxplot", {
  plot <- summarize_data(gapminder, continent, lifeExp)
  expect_silent(print(plot[[2]]))})

```


**I used 5 `expect_()` functions to test the `summarize_data` function and the tests passed.** 

**First,** I computed summary statistics manually, extracted the mean. I also computed summary statistics using the `summarize_data` function and extracted the mean. I then used `expect_identical()` to test if the two means are identical.

**Second,** I used `expect_type()` to test if the `summarize_data` function returns a `list`.

**Third,** I used `expect_error()` to test if using a numeric variable instead of a categorical variable in the `summarize_data` function would pass. The test passed because `expect_error()` is expecting an error.

**Fourth,** I computed summary statistics manually, extracted the mean vector, this time, not removing `NAs`. I also used the `summarize_data` function to compute summary statistics, with `na.rm = FALSE`, and then extracted the mean vector. I used `expect_equal()` to test if the two mean vectors have `NAs`.

**Fifth,** I used `expect_silent()` to test if the `summarize_data` function successfully creates a boxplot.


**References**

1. https://www.youtube.com/watch?v=3nDgR7l5Tps
2. http://adv-r.had.co.nz/Computing-on-the-language.html
3. https://stackoverflow.com/questions/53761025/
   testthatexpect-silent-does-not-seem-to-notice-ggplot2-errors