---
title: "Assignment B-1"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## 1. Introduction

This mini-project is part of the assignments in the STAT545B course. It involves the following exercises:

1. Making a function and fortifying it 
2. Documenting the function 
3. Demonstrating the usefulness of the function with examples
4. Testing the function
5. Making a README file for the project repository

Ultimately, a well written, reproducible Rmd file containing all the code used in the exercises and relevant descriptions will be knitted to a markdown file and pushed to the new Github repository created as part of the course.


### 1.1. Load required packages

The packages below will be used to complete the exercises in this project. Also, the
`palmerpenguins::penguins`, `gapminder::gapminder`, `datateachr::apt_buildings` datasets will be used. Therefore, I will load them before I continue with the exercises. 

However, if any of the packages is not installed, the function `install.packages()` should be used for that purpose before loading.

```{r}
suppressMessages(library(tidyverse))
suppressMessages(library(palmerpenguins))
suppressMessages(library(gapminder))
suppressMessages(library(testthat))
suppressMessages(library(datateachr))
suppressMessages(library(roxygen2))
```
